---
title: "logistic_regression_vs_LDA"
author: "Collin"
date: "11/21/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(tidyverse)
library(ggplot2)
library(MASS)
library(class)
```

## **Homework 5**

1. *Consider Fisher’s iris data set. Extract the data corresponding to the flower types versicolor and virginica, numbering a total of 100 flowers. Set aside the first 10 observations for each flower type as test data and use the remaining data consisting of 80 observations (with flower types as class labels) as training data (1.5 points).*
```{r}
newIris <- iris %>% 
  filter(Species!='setosa')%>% 
  mutate(Species=factor(Species)) 


test<- newIris %>% group_by(Species) %>% slice(1:10) 
train <- newIris %>% group_by(Species) %>% slice(11:50) 



# versicolor <- 1; virginica <-0
#test$Species<-ifelse(test$Species=="versicolor",1,0)
test$Species

#train$Species=ifelse(train$Species == "versicolor",1,0)
  
#now i need training data



```



**2.** *Use LDA for classifying the test data. Use Sepal.Length and Sepal.Width as the predictor variables (or features) (1.5 points).*

```{r}
#2
#**2.** *Use LDA for classifying the test data. Use Sepal.Length and Sepal.Width as the 
#predictor variables (or features) (1.5 points).*
ldaFit <- lda(Species~ Sepal.Length + Sepal.Width,data = train)

```


**2a.** *Report the class-specific means of the predictor variables for the training data (1.5 points).*
```{r}
#**2a.** *Report the class-specific means of the predictor variables for the training data (1.5 points).*
ldaFit$means

```


**2b.** *Compute the confusion matrix for the test data and the misclassification error rate (1.5 points).*
```{r}
#**2b.** *Compute the confusion matrix for the test data and the misclassification error 
#rate (1.5 points).*
# confusion matrix
irisPred <- predict(ldaFit, newdata=test) #use the test data in the model that was trained
                                            #using a different dataset

confusionMatrix <- table(irisPred$class,test$Species,dnn = c('Predicted Species','Species'))
#note the arguments in the table function:
#first argument is the predicted classes from the predict() output
#second argument 
knitr::kable(confusionMatrix)
confusionMatrix

accLDA = sum(diag(confusionMatrix))/length(test$Species)
misclassificationRateLDA = 1 - accLDA
misclassificationRateLDA


ldaVec = c(accLDA,misclassificationRateLDA)
ldaVec

```

**3.** *Use the logistic model, fitted to the training data, to classify the test data.*
```{r}
#logistic regression
#The logistic regression coefficients give the change in the log odds of the outcome 
#for a one unit increase in the predictor variable.

logisticFit = glm(Species ~ Sepal.Length + Sepal.Width, data = train, family = binomial)
#use the train dataset to build the model

```

**3a** *Fit a logistic model to the training data, using the variables Sepal.Length and Sepal.Width as predictors (3 points).*
```{r}
#logistic regression
#The logistic regression coefficients give the change in the log odds of the outcome 
#for a one unit increase in the predictor variable.

logisticFit = glm(Species ~ Sepal.Length + Sepal.Width, data = train, family = binomial)
#use the train dataset to build the model



```



**3ai** *Obtain the estimates and their standard errors for the model parameters (1.5 points).*
```{r}
summary(logisticFit) #estimates of the parameters and std. errors are seen in the table

```


**3aii** *Compute the confusion matrix for the test data and the misclassification error rate (1.5 points).*
```{r}
#ii Compute the confusion matrix for the test data and the misclassification error rate 

# predicted classes
predictedProbs<-predict(logisticFit, type = "response",newdata = test)

predicted<-ifelse(predict(logisticFit, type = "response",newdata = test)>.5,"versicolor","virginica")


#append the test dataset with the predicted values

confMat = table(predicted,test$Species,dnn = c("Predicted Species", "True Species") )
confMat


#misclassification rate = 1 - accuracy
#accuracy
acc = sum(diag(confMat))/length(predicted)
acc
misclassificationRate = 1 - acc
misclassificationRate

logistic2PVec = c(acc,misclassificationRate)
logistic2PVec



```


**3aiii** *Are both of the predictor variables necessary for the purpose of classification? Explain your answer (3 points).*
```{r}
#iii
#Are both of the predictor variables necessary for the purpose of classification? 
summary(logisticFit)
 


```
no, sepal.width is not statistically significant.

**3b** *Fit a logistic regression model to the training data, using the variable Sepal.Length as a one-dimensional predictor (3 points*
```{r}
#b)Fit a logistic regression model to the training data, using the variable 
#Sepal.Length as a one-dimensional predictor (3 points).
logFit2 = glm(Species ~ Sepal.Length, data = train, family = binomial)

```

**3bi** *Obtain the estimates and their standard errors for the model parameters (1.5 points).*
```{r}
#i
summary(logFit2)



```


**3bii** *Compute the confusion matrix for the test data, and the misclassification error rate (1.5 points).*
```{r}
#ii Compute the confusion matrix for the test data, and the misclassification error rate (1.5 points).
predicted2<-ifelse(predict(logFit2, type = "response",newdata = test)>.5,"versicolor","virginica")
confusMat = table(predicted2,test$Species,dnn = c("Predicted Species", "True Species") )
confusMat

acc2 = sum(diag(confusMat))/length(predicted2)
acc2
misclassificationRate2 = 1 - acc2
misclassificationRate2

logistic1PVec = c(acc2,misclassificationRate2)
#logistic1PVec

```
The misclassification rate is .6, or 60%.

**3biii** *Compare the results with those in 3(a). Does your result in 3(b)(ii) support the answer to 3(a)(iii) (3 points)?a*
```{r}
#iii

```
yes, my findings support my answer to 3aiii. Notice the misclassification rates are the 
same, and the Residual Deviance is about the same as well.



**4.** *Use the k-nearest neighbors (k-NN) classification method to classify the test data, using only Sepal.Length as the predictor variable. Perform this analysis using k=1 and k=5. In each case, compute the confusion matrix for the test data, and the misclassification error rate (3 points).*
````{r}

trainSepalLength = train[1:80,1]

testSepalLength = test[1:20,1]
trueClassifications = train[1:80,5]



train=as_tibble(train)
SLTrain = train %>% 
  dplyr::select(Sepal.Length,Species)


SLTest = test %>% 
  dplyr::select(Sepal.Length,Species)




#train and test arguments must be dataframes, but the cl argument (the true class levels of the training data)
#must be a factor vector
knnPredicted1 = knn(train = SLTrain[,1,drop = FALSE],test = SLTest[,1,drop=FALSE], cl = SLTrain$Species, k =1)
knnPredicted1

knnPredicted5 = knn(train = SLTrain[,1,drop = FALSE],test = SLTest[,1,drop=FALSE], cl = SLTrain$Species, k =5)
knnPredicted5


test = test %>% add_column(knnPredicted5)


#for k = 1
confusionMatrixK1 = table(knnPredicted1,test$Species,dnn = c("Predicted Species, k = 1", "Actual"))
confusionMatrixK1


accuracy1 = sum(diag(confusionMatrixK1))/length(knnPredicted1)
accuracy1
mcr1= 1 - accuracy1
mcr1

kNN1Vec = c(accuracy1,mcr1)
kNN1Vec

#for k = 5
confusionMatrixK5 = table(knnPredicted5,test$Species,dnn = c("Predicted Species, k = 5", "Actual"))
confusionMatrixK5

accuracy2 = sum(diag(confusionMatrixK5))/length(knnPredicted5)
accuracy2
mcr2= 1 - accuracy2
mcr2

kNN5Vec = c(accuracy2,mcr2)
kNN5Vec


````


**5.** *Write a very brief summary (maximum of 100 words) about the comparative performance of the three different classification methods for this data set (3 points).*
```{r}
#5 Summary comparing their performances
ldaVec
logistic1PVec
logistic2PVec
kNN1Vec
kNN5Vec


summaryDF = data.frame(LDA = ldaVec,logistic_2_predictor = logistic2PVec,
                       logistic_1_predictor = logistic1PVec,kNN_k_equals_1 = kNN1Vec,
                       kNN_k_equals_5 = kNN5Vec)


rownames(summaryDF) = c("accuracy rate","misclassification rate")
summaryDF


```
Here is a dataframe that summarizes the effectiveness of these 5 different models.
the Linear Discriminant Model and kNN model (where k = 5) performed similarly in that
they both had the same accuracy and misclassification rates. 

Speaking of the kNN model (where k = 5), notice how it actually performed worse than
the kNN model where k = 1. This is good evidence as to why a higher k isn't always better,
even though intuitively at first it may seem like a larger k would be better. The kNN model
where k = 1, actually was the most accurate model, and also had the lowest misclassification rate.

Both logistic models (with one and 2 predictors) performed the same in terms of accuracy 
and their misclassification rates.

